---
title: "01. Data Visualizations"
author: "Nicolas L. Camacho"
date: "2025-08-27"
output: html_document
---

# Install packages (where necessary) and load libraries to run this script
```{r, echo = TRUE, message = FALSE, warning = FALSE}
required_packages <- c("knitr", "tidyverse", "report", "ggplot2")
installed <- required_packages %in% rownames(installed.packages())
if(any(!installed)) {
  install.packages(required_packages[!installed])
}
invisible(lapply(required_packages, library, character.only = T))
```

# Set up the root directory from which the cleaned dataset is pulled
```{r setup, echo = TRUE, message = FALSE, warning = FALSE}
opts_knit$set(root.dir = "C:/Users/nicol/Box/01. Nicolas Dissertation/Data/")
# Update this section to reflect the file path that contains the cleaned dataset
```

# Load cleaned dataset
```{r, echo = TRUE, message = FALSE, warning = FALSE}
trunc_data <- read_csv("truncated_dataset.csv") %>% 
  mutate(across(sex:income, as.factor),
         across(ksads_mdd_anhed:ptsd_cons, as.factor))
```

# Data Visualizations

## Demographic data

### Age

#### Full dataset
```{r}
trunc_data %>% 
  dplyr::select(., age_mos) %>%
  report(., centrality = "mean")
```

#### Per study
```{r}
trunc_data %>%
  dplyr::select(., subid, age_mos) %>%
  mutate(study = case_when(str_detect(subid, "noa") ~ "noa",
                           str_detect(subid, "pcit") ~ "pcit")) %>% 
  group_by(study) %>% 
  report(., centrality = "mean")
```

### All other demographics

### Full dataset
```{r}
demos_summary <- trunc_data %>% 
  dplyr::select(., sex:income) %>%
  report(.)

summary(as.data.frame(demos_summary))
```

### PCIT-ED Study
```{r}
demos_pcit_summary <- trunc_data %>% 
  filter(., str_detect(subid, "pcit")) %>% 
  dplyr::select(., sex:income) %>%
  report(.)

summary(as.data.frame(demos_pcit_summary))
```

### NOA Study
```{r}
demos_noa_summary <- trunc_data %>% 
  filter(., str_detect(subid, "noa")) %>% 
  dplyr::select(., sex:income) %>%
  report(.)

summary(as.data.frame(demos_noa_summary))
```

## Item-level data on the PFC-S per study
```{r, fig.width = 8, fig.height = 10}
level_order <- c("PCIT-ED", "NOA")
trunc_data %>% 
  dplyr::select(., subid, contains("pfc")) %>%
  mutate(across(pfc01:pfc23, factor)) %>% 
  mutate(study = case_when(str_detect(subid, "noa") ~ "NOA",
                           str_detect(subid, "pcit") ~ "PCIT-ED")) %>% 
  pivot_longer(., !c(subid, study), names_to = "vars", values_to = "vals") %>%
  drop_na() %>% 
  mutate(vars = str_remove_all(vars, "pfc")) %>% 
  ggplot(., aes(x = vals, fill = factor(study, level_order))) +
  geom_bar(position = "dodge") +
  scale_fill_manual(values = c("NOA" = "#1E88E5", "PCIT-ED" = "#FFC107")) +
  scale_y_continuous(breaks = seq(0, 200, 25)) +
  facet_wrap(~vars, scales = "fixed") +
  theme_gray(base_family = "serif", base_size = 12) +
  theme(legend.position = "bottom") +
  labs(x = "Response Options",
       y = "Count",
       fill = NULL)
```

## ERC, BIS/BAS, BDI-II, My Child 2, BRIEF GEC, and CBCL (sub)scale data per study
```{r, warning = FALSE}
trunc_data %>%
  dplyr::select(., subid, contains("erc"), bis_mean, contains("bas_"),
                bdi_mean, mca_gltfeel_mean, brief_gec_t_comb, contains("cbcl")) %>% 
  pivot_longer(., !subid, values_to = "vals", names_to = "vars") %>% 
  mutate(study = factor(case_when(str_detect(subid, "pcit") ~ "PCIT-ED",
                                  str_detect(subid, "noa") ~ "NOA"))) %>% 
  ggplot(., aes(x = vals, fill = study)) +
  geom_density(alpha = 0.5) +
  scale_fill_manual(values = c("NOA" = "#1E88E5", "PCIT-ED" = "#FFC107")) +
  facet_wrap(~vars, scales = "free")
```

## KSADS-EC anhedonia item and diagnostic data

### Full dataset
```{r}
trunc_data %>% 
  dplyr::select(., subid, ksads_mdd_anhed, mdd_dx, contains("cons")) %>% 
  mutate(study = factor(case_when(str_detect(subid, "pcit") ~ "PCIT-ED",
                                  str_detect(subid, "noa") ~ "NOA"))) %>%
  pivot_longer(., !c("subid", "study"), names_to = "vars", values_to = "designation") %>% 
  ggplot(., aes(x = designation, fill = study)) +
  geom_bar(position = "dodge") + 
  ylab("Count") +
  xlab("Designation") +
  scale_fill_manual(values = c("NOA" = "#1E88E5", "PCIT-ED" = "#FFC107")) +
  facet_wrap(~vars, scales = "free")
```

# Missing Data

## Update dataset to specify NA = 1 and all other values = 0
```{r}
trunc_2 <- trunc_data %>% 
  dplyr::select(., subid, pfc01:cbcl_tot_t, ksads_mdd_anhed:ptsd_cons) %>% 
  mutate(across(pfc01:ptsd_cons, ~if_else(is.na(.), 1, 0)))
```

## Show missingness counts and percentages per study
```{r}
trunc_2 %>% 
    mutate(across(pfc01:ptsd_cons, factor)) %>% 
    mutate(study = case_when(str_detect(subid, "noa") ~ "noa",
                             str_detect(subid, "pcit") ~ "pcit")) %>% 
  group_by(study) %>% 
  report(.)
```







