---
title: "02. Multiple Imputation"
author: "Nicolas L. Camacho"
date: "2025-08-27"
output: html_document
---

# Install packages (where necessary) and load libraries to run this script
```{r, echo = TRUE, message = FALSE, warning = FALSE}
required_packages <- c("knitr", "tidyverse", "VIM", "naniar", "smd", "stats",
                       "car", "broom", "mice", "miceadds")
installed <- required_packages %in% rownames(installed.packages())
if(any(!installed)) {
  install.packages(required_packages[!installed])
}
invisible(lapply(required_packages, library, character.only = T))
```

# Set up the root directory from which the cleaned dataset is pulled
```{r setup, echo = TRUE, message = FALSE, warning = FALSE}
opts_knit$set(root.dir = "C:/Users/nicol/Box/01. Nicolas Dissertation/Data/")
# Update this section to reflect the file path that contains the cleaned dataset
```

# Load cleaned dataset
```{r, echo = TRUE, message = FALSE, warning = FALSE}
trunc_data <- read_csv("truncated_dataset.csv") %>% 
  mutate(subid = as.factor(subid),
         across(sex:pfc23, as.factor),
         across(ksads_mdd_anhed:ptsd_cons, as.factor))
```

# Add auxiliary variables based on participation in the NOA/PCIT-ED study
## Note: CGAS is already in the dataset and was also used as an auxiliary var
```{r}
trunc_data_2 <- trunc_data %>% 
  mutate(noa_v_pcit_sub = factor(if_else(str_detect(subid, "noa"), 1, 0)), # NOA = 1, PCIT-ED = 0
         pcit_hc_sub = factor(if_else(str_detect(subid, "pcited_hc"), 1, 0)), # PCIT-ED Healthy Controls
         pcit_cc_sub = factor(if_else(str_detect(subid, "pcited_cc"), 1, 0)), # PCIT-ED Community Controls
         .after = income) 
```

# Assess the percent of missing data

## Across the entire dataframe
```{r}
kable(trunc_data_2 %>%
        summarise(
          entire = pct_miss(dplyr::select(., pfc01:ptsd_cons)),
          pfcs = pct_miss(dplyr::select(., contains("pfc"))),
          bdi = pct_miss(dplyr::select(., bdi_mean)),
          bis_bas = pct_miss(dplyr::select(., bis_mean:bas_reward_mean)),
          erc = pct_miss(dplyr::select(., contains("erc"))),
          mc2 = pct_miss(dplyr::select(., mca_gltfeel_mean)),
          brief = pct_miss(dplyr::select(., brief_gec_t_comb)),
          cbcl = pct_miss(dplyr::select(., contains("cbcl"))),
          cgas = pct_miss(dplyr::select(., cgas)),
          ksads_mdd_anhed = pct_miss(dplyr::select(., ksads_mdd_anhed)),
          ksads_dx = pct_miss(dplyr::select(., mdd_dx, contains("cons")))))
```

## NOA
```{r}
kable(trunc_data_2 %>% 
        filter(str_detect(subid, "noa")) %>%
        summarise(
          entire = pct_miss(dplyr::select(., pfc01:ptsd_cons)),
          pfcs = pct_miss(dplyr::select(., contains("pfc"))),          
          bdi = pct_miss(dplyr::select(., bdi_mean)),
          bis_bas = pct_miss(dplyr::select(., bis_mean:bas_reward_mean)),
          erc = pct_miss(dplyr::select(., contains("erc"))),
          mc2 = pct_miss(dplyr::select(., mca_gltfeel_mean)),
          brief = pct_miss(dplyr::select(., brief_gec_t_comb)),
          cbcl = pct_miss(dplyr::select(., contains("cbcl"))),
          cgas = pct_miss(dplyr::select(., cgas)),
          ksads_mdd_anhed = pct_miss(dplyr::select(., ksads_mdd_anhed)),
          ksads_dx = pct_miss(dplyr::select(., mdd_dx, contains("cons")))))
```


## PCIT-ED
```{r}
kable(trunc_data_2 %>% 
        filter(str_detect(subid, "pcit")) %>%
        summarise(
          entire = pct_miss(dplyr::select(., pfc01:ptsd_cons)),
          pfcs = pct_miss(dplyr::select(., contains("pfc"))),
          bdi = pct_miss(dplyr::select(., bdi_mean)),
          bis_bas = pct_miss(dplyr::select(., bis_mean:bas_reward_mean)),
          erc = pct_miss(dplyr::select(., contains("erc"))),
          mc2 = pct_miss(dplyr::select(., mca_gltfeel_mean)),
          brief = pct_miss(dplyr::select(., brief_gec_t_comb)),
          cbcl = pct_miss(dplyr::select(., contains("cbcl"))),
          cgas = pct_miss(dplyr::select(., cgas)),
          ksads_mdd_anhed = pct_miss(dplyr::select(., ksads_mdd_anhed)),
          ksads_dx = pct_miss(dplyr::select(., mdd_dx, contains("cons")))))
```

# Create dummy variables for all variables with missing data
# and remove dummy variables that will not be imputed (e.g., child education)
```{r}
miss_dum_fun <- list(miss_dum = ~factor(if_else(is.na(.x), 1, 0)))

trunc_data_3 <- trunc_data_2 %>% 
  mutate(across(where(any_na), miss_dum_fun, .names = "{.col}_{.fn}")) %>% 
  dplyr::select(., -racespec_miss_dum, -child_edu_miss_dum, -informant_miss_dum,
                -income_miss_dum, -ksads_mdd_anhed_miss_dum, -mdd_dx_miss_dum, 
                -contains("cons_miss_dum"))
```

# Test which variables in the dataset are useful predictors of the missing data

## PFC-S missing data (very few)

### Using Standardized Mean Differences for continuous variables
### Cohen's effect size of +/- .20 (small effect) is minimum threshold
```{r}
pfcs_smd_summary <- trunc_data_3 %>% 
  pivot_longer(., c(pfc02_miss_dum:pfc22_miss_dum),
               names_to = "pfc_miss_dum_vars", values_to = "vals") %>% 
  group_by(pfc_miss_dum_vars) %>% 
  summarise(across(c(age_mos, bdi_mean, erc_er_mean, erc_ln_mean, 
                     brief_gec_t_comb), ~smd(x = .x, g = vals, na.rm = T)$estimate)) %>% 
  rename(., vars = pfc_miss_dum_vars) %>% 
  mutate(., vars = str_remove(vars, "_miss_dum"))
kable(pfcs_smd_summary)

# Other continuous variables could not be used as they had missing data in cases
# where data was missing for the PFC-S items
```

### Binary summary of which continuous variables should be predictors of missing data
```{r}
pfcs_smd_binary <- pfcs_smd_summary %>% 
  mutate(across(where(is.numeric), ~if_else(abs(.x) >= .20, 1, 0)))

kable(pfcs_smd_binary)
```

### Logistic regression p-value results for categorical variables
```{r, warning = F}
pfcs_logreg_summary <- trunc_data_3 %>% 
  dplyr::select(., sex, race, hispanic, child_edu, informant, income, 
                pfc01:pfc23, noa_v_pcit_sub, pcit_hc_sub, pcit_cc_sub, 
                ksads_mdd_anhed, mdd_dx, contains("_cons"), 
                pfc02_miss_dum:pfc22_miss_dum) %>%
  pivot_longer(., c(pfc02_miss_dum:pfc22_miss_dum),
               names_to = "pfcs_miss_dum_vars", values_to = "vals") %>% 
  group_by(pfcs_miss_dum_vars) %>% 
  summarise(across(c(sex:ptsd_cons), ~tidy(Anova(glm(vals ~ 1 + .x,
                                                       family = "binomial", data = .),
                                                   type = "III"))$p.value)) %>% 
  rename(., vars = pfcs_miss_dum_vars) %>% 
  mutate(., vars = str_remove(vars, "_miss_dum"))

kable(pfcs_logreg_summary)
```

### Binary summary of which categorical variables should be predictors of missing data
```{r}
pfcs_logreg_binary <- pfcs_logreg_summary %>% 
  mutate(across(where(is.numeric), ~if_else(.x < .05, 1, 0)))

kable(pfcs_logreg_binary)
```

## BIS/BAS missing data (for PCIT-ED participants only)

### Using Standardized Mean Differences for continuous variables
### Cohen's effect size of +/- .20 (small effect) is minimum threshold
```{r}
bisbas_smd_summary <- trunc_data_3 %>% 
  filter(str_detect(subid, "pcit")) %>% 
  pivot_longer(., c(bis_mean_miss_dum:bas_reward_mean_miss_dum),
               names_to = "bisbas_miss_dum_vars", values_to = "vals") %>% 
  group_by(bisbas_miss_dum_vars) %>% 
  summarise(across(c(age_mos, bdi_mean, erc_er_mean, erc_ln_mean,
                     mca_gltfeel_mean, brief_gec_t_comb, cbcl_int_t, cbcl_ext_t,
                     cbcl_tot_t, cgas), ~smd(x = .x, g = vals, na.rm = T)$estimate)) %>% 
  rename(., vars = bisbas_miss_dum_vars) %>% 
  mutate(., vars = str_remove(vars, "_miss_dum"))
kable(bisbas_smd_summary)

# Other continuous variables could not be used as they had missing data in cases
# where data was missing for the PFC-S items
```

### Binary summary of which continuous variables should be predictors of missing data
```{r}
bisbas_smd_binary <- bisbas_smd_summary %>% 
  mutate(across(where(is.numeric), ~if_else(abs(.x) >= .20, 1, 0)))

kable(bisbas_smd_binary)
```

### Logistic regression p-value results for categorical variables
```{r, warning = F}
bisbas_logreg_summary <- trunc_data_3 %>% 
  filter(str_detect(subid, "pcit")) %>%
  dplyr::select(., sex, race, hispanic, child_edu, informant, income, 
                pfc01:pfc23, pcit_hc_sub, pcit_cc_sub, ksads_mdd_anhed, mdd_dx,
                contains("_cons"), bis_mean_miss_dum:bas_reward_mean_miss_dum) %>%
  pivot_longer(., c(bis_mean_miss_dum:bas_reward_mean_miss_dum),
               names_to = "bisbas_miss_dum_vars", values_to = "vals") %>% 
  group_by(bisbas_miss_dum_vars) %>% 
  summarise(across(c(sex:ptsd_cons), ~tidy(Anova(glm(vals ~ 1 + .x,
                                                       family = "binomial", data = .),
                                                   type = "III"))$p.value)) %>% 
  rename(., vars = bisbas_miss_dum_vars) %>% 
  mutate(., vars = str_remove(vars, "_miss_dum"))

kable(bisbas_logreg_summary)
```

### Binary summary of which categorical variables should be predictors of missing data
```{r}
bisbas_logreg_binary <- bisbas_logreg_summary %>% 
  mutate(across(where(is.numeric), ~if_else(.x < .05, 1, 0)))

kable(bisbas_logreg_binary)
```

## ERC missing data

### Using Standardized Mean Differences for continuous variables
### Cohen's effect size of +/- .20 (small effect) is minimum threshold
```{r}
erc_smd_summary <- trunc_data_3 %>% 
  pivot_longer(., c(erc_er_mean_miss_dum, erc_ln_mean_miss_dum),
               names_to = "erc_miss_dum_vars", values_to = "vals") %>% 
  group_by(erc_miss_dum_vars) %>% 
  summarise(across(c(age_mos, bdi_mean, bis_mean, bas_mean, bas_drive_mean,
                     bas_funseek_mean, bas_reward_mean, brief_gec_t_comb, 
                     cbcl_int_t, cbcl_ext_t, cbcl_tot_t, cgas), 
                   ~smd(x = .x, g = vals, na.rm = T)$estimate)) %>% 
  rename(., vars = erc_miss_dum_vars) %>% 
  mutate(., vars = str_remove(vars, "_miss_dum"))
kable(erc_smd_summary)

# Other continuous variables could not be used as they had missing data in cases
# where data was missing for the PFC-S items
```

### Binary summary of which continuous variables should be predictors of missing data
```{r}
erc_smd_binary <- erc_smd_summary %>% 
  mutate(across(where(is.numeric), ~if_else(abs(.x) >= .20, 1, 0)))

kable(erc_smd_binary)
```

### Logistic regression p-value results for categorical variables
```{r, warning = F}
erc_logreg_summary <- trunc_data_3 %>% 
  dplyr::select(., sex, race, hispanic, child_edu, informant, income, 
                pfc01:pfc23, pcit_hc_sub, pcit_cc_sub, ksads_mdd_anhed, mdd_dx,
                contains("_cons"), erc_er_mean_miss_dum, erc_ln_mean_miss_dum) %>%
  pivot_longer(., c(erc_er_mean_miss_dum, erc_ln_mean_miss_dum),
               names_to = "erc_miss_dum_vars", values_to = "vals") %>% 
  group_by(erc_miss_dum_vars) %>% 
  summarise(across(c(sex:ptsd_cons), ~tidy(Anova(glm(vals ~ 1 + .x,
                                                       family = "binomial", data = .),
                                                   type = "III"))$p.value)) %>% 
  rename(., vars = erc_miss_dum_vars) %>% 
  mutate(., vars = str_remove(vars, "_miss_dum"))

kable(erc_logreg_summary)
```

### Binary summary of which categorical variables should be predictors of missing data
```{r}
erc_logreg_binary <- erc_logreg_summary %>% 
  mutate(across(where(is.numeric), ~if_else(.x < .05, 1, 0)))

kable(erc_logreg_binary)
```

## My Child 2 missing data

### Using Standardized Mean Differences for continuous variables
### Cohen's effect size of +/- .20 (small effect) is minimum threshold
```{r}
mca_smd_summary <- trunc_data_3 %>% 
  filter(str_detect(subid, "pcit")) %>% 
  pivot_longer(., c(mca_gltfeel_mean_miss_dum),
               names_to = "mca_miss_dum_vars", values_to = "vals") %>% 
  group_by(mca_miss_dum_vars) %>% 
  summarise(across(c(age_mos, bdi_mean, bis_mean, bas_mean, bas_drive_mean,
                     bas_funseek_mean, bas_reward_mean, erc_er_mean, erc_ln_mean,
                     brief_gec_t_comb, cbcl_int_t, cbcl_ext_t, cbcl_tot_t, cgas), 
                   ~smd(x = .x, g = vals, na.rm = T)$estimate)) %>% 
  rename(., vars = mca_miss_dum_vars) %>% 
  mutate(., vars = str_remove(vars, "_miss_dum"))
kable(mca_smd_summary)

# Other continuous variables could not be used as they had missing data in cases
# where data was missing for the PFC-S items
```

### Binary summary of which continuous variables should be predictors of missing data
```{r}
mca_smd_binary <- mca_smd_summary %>% 
  mutate(across(where(is.numeric), ~if_else(abs(.x) >= .20, 1, 0)))

kable(mca_smd_binary)
```

### Logistic regression p-value results for categorical variables
```{r, warning = F}
mca_logreg_summary <- trunc_data_3 %>% 
  filter(str_detect(subid, "pcit")) %>% 
  dplyr::select(., sex, race, hispanic, child_edu, informant, income, 
                pfc01:pfc23, pcit_hc_sub, pcit_cc_sub, ksads_mdd_anhed, mdd_dx,
                contains("_cons"), mca_gltfeel_mean_miss_dum) %>%
  pivot_longer(., mca_gltfeel_mean_miss_dum,
               names_to = "mca_miss_dum_vars", values_to = "vals") %>% 
  group_by(mca_miss_dum_vars) %>% 
  summarise(across(c(sex:ptsd_cons), ~tidy(Anova(glm(vals ~ 1 + .x,
                                                       family = "binomial", data = .),
                                                   type = "III"))$p.value)) %>% 
  rename(., vars = mca_miss_dum_vars) %>% 
  mutate(., vars = str_remove(vars, "_miss_dum"))

kable(mca_logreg_summary)
```

### Binary summary of which categorical variables should be predictors of missing data
```{r}
mca_logreg_binary <- mca_logreg_summary %>% 
  mutate(across(where(is.numeric), ~if_else(.x < .05, 1, 0)))

kable(mca_logreg_binary)
```

## BRIEF GEC missing data

### Using Standardized Mean Differences for continuous variables
### Cohen's effect size of +/- .20 (small effect) is minimum threshold
```{r}
brief_smd_summary <- trunc_data_3 %>% 
  pivot_longer(., brief_gec_t_comb_miss_dum,
               names_to = "brief_miss_dum_vars", values_to = "vals") %>% 
  group_by(brief_miss_dum_vars) %>% 
  summarise(across(c(age_mos, bdi_mean, bis_mean, bas_mean, bas_drive_mean,
                     bas_funseek_mean, bas_reward_mean, erc_er_mean, erc_ln_mean,
                     mca_gltfeel_mean, cbcl_int_t, cbcl_ext_t,
                     cbcl_tot_t, cgas), 
                   ~smd(x = .x, g = vals, na.rm = T)$estimate)) %>% 
  rename(., vars = brief_miss_dum_vars) %>% 
  mutate(., vars = str_remove(vars, "_miss_dum"))
kable(brief_smd_summary)
```

### Binary summary of which continuous variables should be predictors of missing data
```{r}
brief_smd_binary <- brief_smd_summary %>% 
  mutate(across(where(is.numeric), ~if_else(abs(.x) >= .20, 1, 0)))

kable(brief_smd_binary)
```

### Logistic regression p-value results for categorical variables
```{r, warning = F}
brief_logreg_summary <- trunc_data_3 %>% 
  dplyr::select(., sex, race, hispanic, child_edu, informant, income, 
                pfc01:pfc23, pcit_hc_sub, pcit_cc_sub, ksads_mdd_anhed, mdd_dx,
                contains("_cons"), brief_gec_t_comb_miss_dum) %>%
  pivot_longer(., brief_gec_t_comb_miss_dum,
               names_to = "brief_miss_dum_vars", values_to = "vals") %>% 
  group_by(brief_miss_dum_vars) %>% 
  summarise(across(c(sex:ptsd_cons), ~tidy(Anova(glm(vals ~ 1 + .x,
                                                       family = "binomial", data = .),
                                                   type = "III"))$p.value)) %>% 
  rename(., vars = brief_miss_dum_vars) %>% 
  mutate(., vars = str_remove(vars, "_miss_dum"))

kable(brief_logreg_summary)
```

### Binary summary of which categorical variables should be predictors of missing data
```{r}
brief_logreg_binary <- brief_logreg_summary %>% 
  mutate(across(where(is.numeric), ~if_else(.x < .05, 1, 0)))

kable(brief_logreg_binary)
```

## CBCL missing data

### Using Standardized Mean Differences for continuous variables
### Cohen's effect size of +/- .20 (small effect) is minimum threshold
```{r}
cbcl_smd_summary <- trunc_data_3 %>% 
  pivot_longer(., c(cbcl_int_t_miss_dum:cbcl_tot_t_miss_dum),
               names_to = "cbcl_miss_dum_vars", values_to = "vals") %>% 
  group_by(cbcl_miss_dum_vars) %>% 
  summarise(across(c(age_mos, bdi_mean, bis_mean, bas_mean, bas_drive_mean,
                     bas_funseek_mean, bas_reward_mean, erc_er_mean, erc_ln_mean,
                     mca_gltfeel_mean, brief_gec_t_comb, cgas), 
                   ~smd(x = .x, g = vals, na.rm = T)$estimate)) %>% 
  rename(., vars = cbcl_miss_dum_vars) %>% 
  mutate(., vars = str_remove(vars, "_miss_dum"))
kable(cbcl_smd_summary)
```

### Binary summary of which continuous variables should be predictors of missing data
```{r}
cbcl_smd_binary <- cbcl_smd_summary %>% 
  mutate(across(where(is.numeric), ~if_else(abs(.x) >= .20, 1, 0)))

kable(cbcl_smd_binary)
```

### Logistic regression p-value results for categorical variables
```{r, warning = F}
cbcl_logreg_summary <- trunc_data_3 %>% 
  dplyr::select(., sex, race, hispanic, child_edu, informant, income, 
                pfc01:pfc23, pcit_hc_sub, pcit_cc_sub, ksads_mdd_anhed, mdd_dx,
                contains("_cons"), contains("cbcl")) %>%
  pivot_longer(., c(cbcl_int_t_miss_dum:cbcl_tot_t_miss_dum),
               names_to = "cbcl_miss_dum_vars", values_to = "vals") %>% 
  group_by(cbcl_miss_dum_vars) %>% 
  summarise(across(c(sex:ptsd_cons), ~tidy(Anova(glm(vals ~ 1 + .x,
                                                       family = "binomial", data = .),
                                                   type = "III"))$p.value)) %>% 
  rename(., vars = cbcl_miss_dum_vars) %>% 
  mutate(., vars = str_remove(vars, "_miss_dum"))

kable(cbcl_logreg_summary)
```

### Binary summary of which categorical variables should be predictors of missing data
```{r}
cbcl_logreg_binary <- cbcl_logreg_summary %>% 
  mutate(across(where(is.numeric), ~if_else(.x < .05, 1, 0)))

kable(cbcl_logreg_binary)
```

## CGAS missing data

### Using Standardized Mean Differences for continuous variables
### Cohen's effect size of +/- .20 (small effect) is minimum threshold
```{r}
cgas_smd_summary <- trunc_data_3 %>% 
  pivot_longer(., cgas_miss_dum,
               names_to = "cgas_miss_dum_vars", values_to = "vals") %>% 
  group_by(cgas_miss_dum_vars) %>% 
  summarise(across(c(age_mos, bdi_mean, bis_mean, bas_mean, bas_drive_mean,
                     bas_funseek_mean, bas_reward_mean, erc_er_mean, erc_ln_mean,
                     mca_gltfeel_mean, brief_gec_t_comb, cbcl_int_t, cbcl_ext_t,
                     cbcl_tot_t), 
                   ~smd(x = .x, g = vals, na.rm = T)$estimate)) %>% 
  rename(., vars = cgas_miss_dum_vars) %>% 
  mutate(., vars = str_remove(vars, "_miss_dum"))
kable(cgas_smd_summary)
```

### Binary summary of which continuous variables should be predictors of missing data
```{r}
cgas_smd_binary <- cgas_smd_summary %>% 
  mutate(across(where(is.numeric), ~if_else(abs(.x) >= .20, 1, 0)))

kable(cgas_smd_binary)
```

### Logistic regression p-value results for categorical variables
```{r, warning = F}
cgas_logreg_summary <- trunc_data_3 %>% 
  dplyr::select(., sex, race, hispanic, child_edu, informant, income, 
                pfc01:pfc23, pcit_hc_sub, pcit_cc_sub, ksads_mdd_anhed, mdd_dx,
                contains("_cons"), cgas_miss_dum) %>%
  pivot_longer(., cgas_miss_dum,
               names_to = "cgas_miss_dum_vars", values_to = "vals") %>% 
  group_by(cgas_miss_dum_vars) %>% 
  summarise(across(c(sex:ptsd_cons), ~tidy(Anova(glm(vals ~ 1 + .x,
                                                       family = "binomial", data = .),
                                                   type = "III"))$p.value)) %>% 
  rename(., vars = cgas_miss_dum_vars) %>% 
  mutate(., vars = str_remove(vars, "_miss_dum"))

kable(cgas_logreg_summary)
```

### Binary summary of which categorical variables should be predictors of missing data
```{r}
cgas_logreg_binary <- cgas_logreg_summary %>% 
  mutate(across(where(is.numeric), ~if_else(.x < .05, 1, 0)))

kable(cgas_logreg_binary)
```

# Begin multiple imputation process

## Start by creating an imputation instance with no iterations (to access predictor and imputation method matrices)
```{r}
imp0 <- mice(trunc_data_2, maxit = 0)
```

## Extract predictor and imputation method matrices
```{r}
predMat <- imp0$predictorMatrix
impMeth <- imp0$method
```

## Set up the predictor matrix according to results of SMD and logistic regression analyses
```{r}
predMat_res <- data.frame(predMat) %>%
  rownames_to_column(., "vars") %>% 
  mutate(across(-vars, ~case_when(.x >= 0 ~ 0))) %>% 
  
  # for PFC-S items
  rows_update(., data.frame(pfcs_smd_binary), by = "vars") %>% 
  rows_update(., data.frame(pfcs_logreg_binary), by = "vars") %>%
  # for BIS/BAS subscales
  rows_update(., data.frame(bisbas_smd_binary), by = "vars") %>% 
  rows_update(., data.frame(bisbas_logreg_binary), by = "vars") %>%
  # for ERC subscales
  rows_update(., data.frame(erc_smd_binary), by = "vars") %>% 
  rows_update(., data.frame(erc_logreg_binary), by = "vars") %>%
  # for MCA subscale
  rows_update(., data.frame(mca_smd_binary), by = "vars") %>% 
  rows_update(., data.frame(mca_logreg_binary), by = "vars") %>%
  # for BRIEF subscale
  rows_update(., data.frame(brief_smd_binary), by = "vars") %>% 
  rows_update(., data.frame(brief_logreg_binary), by = "vars") %>%
  # for CBCL subscales
  rows_update(., data.frame(cbcl_smd_binary), by = "vars") %>% 
  rows_update(., data.frame(cbcl_logreg_binary), by = "vars") %>%
  # for CGAS 
  rows_update(., data.frame(cgas_smd_binary), by = "vars") %>% 
  rows_update(., data.frame(cgas_logreg_binary), by = "vars") %>%
  
  column_to_rownames(., var = "vars") %>% 
  as.matrix(.)
```

### Visualize the predictor matrix
```{r, fig.height = 8, fig.width = 18}
predMat_res %>% 
  as.data.frame(.) %>% 
  rownames_to_column(., var = "row_vars") %>%
  mutate(row_vars = factor(row_vars, levels = unique(.$row_vars))) %>% 
  pivot_longer(!row_vars, names_to = "predictors", values_to = "vals") %>% 
  mutate(predictors = factor(predictors, levels = unique(.$predictors))) %>% 
  ggplot(data = ., aes(x = predictors, y = row_vars, fill = vals)) +
  geom_tile() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "none")
```

## Set up the method matrix with predictive mean matching and proportional odds model algorithms for continuous and ordinal variables with missing data, respectively
```{r}
impMeth[c(5, 7:9, 50:63)] <- "" # remove imputation of demographics + diagnoses
impMeth[c(14, 16, 18, 20, 26, 30, 32, 34)] <- "polr" # proportional odds model, PFC-S items
impMeth[c(37:49)] <- "pmm" # predictive mean matching model, BIS/BAS, ERC, MC2, BRIEF, CBCL, CGAS
kable(impMeth)
```

# Run multiple imputation (40 datasets) on dataset
# while omitting the imputation of BIS/BAS and My Child 2 data
# for NOA Study participants
```{r}
where_no_noa <- make.where(trunc_data_2)
where_no_noa[1:173, c("bis_mean", "bas_mean", "bas_drive_mean", "bas_funseek_mean",
                      "bas_reward_mean", "mca_gltfeel_mean")] <- FALSE
mi_data <- mice(trunc_data_2, m = 40, method = impMeth,
                predictorMatrix = predMat_res, maxit = 1, printFlag = F,
                where = where_no_noa, seed = 02042025)
```

## Make sure there are no missing values for the variables that were imputed

### Full dataset
```{r}
complete(mi_data, 1) %>% 
  dplyr::select(., pfc01:cgas) %>% 
  aggr(., comb = T, numbers = T, prop = F, cex.axis = 0.5)
```

### NOA Study
```{r}
complete(mi_data, 1) %>% 
  filter(str_detect(subid, "noa")) %>% 
  dplyr::select(., pfc01:cgas) %>% 
  aggr(., comb = T, numbers = T, prop = F, cex.axis = 0.5)
```

### PCIT-ED Study
```{r}
complete(mi_data, 1) %>% 
  filter(str_detect(subid, "noa")) %>% 
  dplyr::select(., pfc01:cgas) %>% 
  aggr(., comb = T, numbers = T, prop = F, cex.axis = 0.5)
```

### Issues arising from common missingness among continuous variables and 
### diagnostic data

```{r}
complete(mi_data, 1) %>% 
  dplyr::select(., sex, race, hispanic:income, bis_mean:ptsd_cons) %>% 
  aggr(., comb = T, numbers = T, prop = F, cex.axis = 0.5)
```

## Assess Logged Events
```{r}
head(mi_data$loggedEvents, n = 11) # same 11 logged events across imputations
```

## Create new predictor matrix without diagnostic variables as predictors
```{r}
predMat_res_2 <- as.data.frame(predMat_res) %>% 
  mutate(across(ksads_mdd_anhed:ptsd_cons, ~0)) %>% 
  as.matrix(.)
```

### Visualize the predictor matrix
```{r, fig.height = 8, fig.width = 18}
predMat_res_2 %>% 
  as.data.frame(.) %>% 
  rownames_to_column(., var = "row_vars") %>%
  mutate(row_vars = factor(row_vars, levels = unique(.$row_vars))) %>% 
  pivot_longer(!row_vars, names_to = "predictors", values_to = "vals") %>% 
  mutate(predictors = factor(predictors, levels = unique(.$predictors))) %>% 
  ggplot(data = ., aes(x = predictors, y = row_vars, fill = vals)) +
  geom_tile() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "none")
```

# Re-run multiple imputation using new predictor matrix (attempt #2)
```{r}
mi_data_2 <- mice(trunc_data_2, m = 40, method = impMeth,
                predictorMatrix = predMat_res_2, maxit = 1, printFlag = F,
                where = where_no_noa, seed = 02042025)
```

## Make sure there are no missing values for the variables that were imputed

### Full dataset
```{r}
complete(mi_data_2, 1) %>% 
  dplyr::select(., pfc01:cgas) %>% 
  aggr(., comb = T, numbers = T, prop = F, cex.axis = 0.5)
```

### NOA Study
```{r}
complete(mi_data_2, 1) %>% 
  filter(str_detect(subid, "noa")) %>% 
  dplyr::select(., sex, race, hispanic:income, pfc01:cgas) %>% 
  aggr(., comb = T, numbers = T, prop = F, cex.axis = 0.5)
```

### PCIT-ED Study
```{r}
complete(mi_data_2, 1) %>% 
  filter(str_detect(subid, "pcit")) %>% 
  dplyr::select(., sex, race, hispanic:income, pfc01:cgas) %>% 
  aggr(., comb = T, numbers = T, prop = F, cex.axis = 0.5)
```

### Issues arising from common missingness between the CGAS and 
### certain variables not uniformly assessed in the PCIT-ED study
### (i.e., child education and family income) and not at all assessed in the
### NOA Study (i.e., BIS/BAS and My Child 2)
```{r}
trunc_data_2 %>% 
  dplyr::select(., sex, race, hispanic:income, bis_mean:cgas) %>% 
  aggr(., comb = T, numbers = T, prop = F, cex.axis = 0.5)
```

## Assess Logged Events
```{r}
head(mi_data_2$loggedEvents, n = 6) # same 6 logged events across imputations
```

## Create new predictor matrix removing child edu and income as predictor of CGAS
## and BIS/BAS and My Child 2 as predictors of the ERC, CBCL, BRIEF, and CGAS
```{r}
predMat_res_3 <- as.data.frame(predMat_res_2) %>% 
  rownames_to_column(., var = "row_names") %>% 
  mutate(
    bis_mean = if_else(row_names == "erc_er_mean" | row_names == "erc_ln_mean" |
                         row_names == "cbcl_int_t" | row_names == "cbcl_ext_t" |
                         row_names == "cbcl_tot_t" | row_names == "brief_gec_t_comb" |
                         row_names == "cgas", 0, bis_mean),
    bas_mean = if_else(row_names == "erc_er_mean" | row_names == "erc_ln_mean" |
                         row_names == "cbcl_int_t" | row_names == "cbcl_ext_t" |
                         row_names == "cbcl_tot_t" | row_names == "brief_gec_t_comb" |
                         row_names == "cgas", 0, bas_mean),
    bas_drive_mean = if_else(row_names == "erc_er_mean" | row_names == "erc_ln_mean" |
                         row_names == "cbcl_int_t" | row_names == "cbcl_ext_t" |
                         row_names == "cbcl_tot_t" | row_names == "brief_gec_t_comb" |
                         row_names == "cgas", 0, bas_drive_mean),
    bas_funseek_mean = if_else(row_names == "erc_er_mean" | row_names == "erc_ln_mean" |
                         row_names == "cbcl_int_t" | row_names == "cbcl_ext_t" |
                         row_names == "cbcl_tot_t" | row_names == "brief_gec_t_comb" |
                         row_names == "cgas", 0, bas_funseek_mean),
    bas_reward_mean = if_else(row_names == "erc_er_mean" | row_names == "erc_ln_mean" |
                         row_names == "cbcl_int_t" | row_names == "cbcl_ext_t" |
                         row_names == "cbcl_tot_t" | row_names == "brief_gec_t_comb" |
                         row_names == "cgas", 0, bas_reward_mean),
    mca_gltfeel_mean = if_else(row_names == "erc_er_mean" | row_names == "erc_ln_mean" |
                         row_names == "cbcl_int_t" | row_names == "cbcl_ext_t" |
                         row_names == "cbcl_tot_t" | row_names == "brief_gec_t_comb" |
                         row_names == "cgas", 0, mca_gltfeel_mean),
    child_edu = if_else(row_names == "cgas", 0, child_edu),
    income = if_else(row_names == "cgas", 0, income)) %>% 
  column_to_rownames(., var = "row_names") %>% 
  as.matrix(.)
```

### Visualize the predictor matrix
```{r, fig.height = 12, fig.width = 18}
pred_order <- c("age_mos", "sex", "race", "racespec", "hispanic", "child_edu",
                "informant", "income", "pfc01", "pfc02", "pfc03", "pfc04",
                "pfc05", "pfc06", "pfc07", "pfc08", "pfc09", "pfc10", "pfc11", 
                "pfc12", "pfc13", "pfc14", "pfc15", "pfc16", "pfc17", "pfc18",
                "pfc19", "pfc20", "pfc21", "pfc22", "pfc23", "bdi", 
                "bis", "bas", "bas_drive", "bas_funseek",
                "bas_reward", "erc_er", "erc_ln", "mca_gltfeel",
                "brief_gec_t", "cbcl_int_t", "cbcl_ext_t", "cbcl_tot_t", 
                "noa_v_pcit_sub", "pcit_hc_sub", "pcit_cc_sub", "cgas")
predMat_res_3 %>% 
  as.data.frame(.) %>% 
  rownames_to_column(., var = "row_vars") %>%
  mutate(row_vars = factor(row_vars, levels = unique(.$row_vars))) %>% 
  pivot_longer(!row_vars, names_to = "predictors", values_to = "vals") %>% 
  mutate(predictors = factor(predictors, levels = unique(.$predictors))) %>% 
  filter(!str_detect(row_vars, "_cons") & !str_detect(predictors, "_cons")) %>% 
  filter(!str_detect(row_vars, "mdd_") & !str_detect(predictors, "mdd_")) %>% 
  filter(row_vars != "subid" & row_vars != "age_mos" & row_vars != "sex" &
           row_vars != "race" & row_vars != "racespec" & row_vars != "hispanic" &
           row_vars != "child_edu" & row_vars != "informant" & row_vars != "income" &
           row_vars != "noa_v_pcit_sub" & row_vars != "pcit_hc_sub" &
           row_vars != "pcit_cc_sub" & row_vars != "cgas" & predictors != "subid") %>%
  mutate(row_vars = str_remove(row_vars, "_mean"),
         row_vars = str_remove(row_vars, "_comb"),
         predictors = str_remove(predictors, "_mean"),
         predictors = str_remove(predictors, "_comb")) %>% 
  ggplot(data = ., aes(x = factor(predictors, pred_order), 
                       y = row_vars, 
                       fill = vals)) +
  geom_tile() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "none",
        text = element_text(family = "serif", size = 24)) +
  labs(x = "Predictors",
       y = "Variables")
```

# Re-run multiple imputation using new predictor matrix
```{r}
mi_data_3 <- mice(trunc_data_2, m = 40, method = impMeth,
                predictorMatrix = predMat_res_3, maxit = 1, printFlag = F,
                where = where_no_noa, seed = 02042025)
```

## Make sure there are no missing values for the variables that were imputed

### Full dataset
```{r}
complete(mi_data_3) %>%  
  dplyr::select(., pfc01:cgas) %>% 
  aggr(., comb = T, numbers = T, prop = F, cex.axis = 0.5)
```

### NOA Study
```{r}
complete(mi_data_3) %>%  
  filter(str_detect(subid, "noa")) %>% 
  dplyr::select(., pfc01:cgas) %>% 
  aggr(., comb = T, numbers = T, prop = F, cex.axis = 0.5)
```

### PCIT-ED Study
```{r}
complete(mi_data_3) %>%  
  filter(str_detect(subid, "pcit")) %>% 
  dplyr::select(., pfc01:cgas) %>% 
  aggr(., comb = T, numbers = T, prop = F, cex.axis = 0.5)
```

## Assess Logged Events
```{r}
head(mi_data_3$loggedEvents, n = 3) # same 3 logged events across imputations
```

# Make sure imputed data is feasible for range of variables
```{r, fig.width = 12, fig.height = 8, message = F}
complete(mi_data_3, "long") %>% 
  mutate(across(pfc01:cgas, as.numeric)) %>% 
  pivot_longer(., c(pfc01:cgas), names_to = "imputed_vars", values_to = "vals") %>% 
  ggplot(aes(x = vals)) + 
  geom_histogram() + 
  facet_wrap(~imputed_vars, scales = "free")
```

# Update the root directory to be the Data folder
```{r}
opts_knit$set(root.dir = "C:/Users/nicol/Box/01. Nicolas Dissertation/Data/")
# Update to reflect the file path in which you would like to store the
# multiply imputed data
```

# Save information and data of the multiple imputations
```{r, include = F}
write.mice.imputation(mi.res = mi_data_3, 
                      name = "multiply_imputed_data", 
                      dattype = "csv",
                      mids2spss = F)
```

# Edit multiply imputed dataset to be Mplus friendly

## For Aims 1 and 2
### Age, sex, all PFC-S items, BIS/BAS, ERC, BDI-II, My Child 2, BRIEF, KSADS Anhedonia
```{r, include = F}
cbind(lapply(lapply(complete(mi_data_3, "long"), is.na), sum))

mi_data_3_mplus <- complete(mi_data_3, "long", include = T) %>% 
  dplyr::select(., .imp, .id, age_mos, sex, pfc01:brief_gec_t_comb, ksads_mdd_anhed) %>% 
  mutate(sex = case_when(sex == "Female" ~ 0, sex == "Male" ~ 1),
         ksads_mdd_anhed = as.numeric(as.character(ksads_mdd_anhed))) %>% 
  replace_na(., list(bis_mean = 999, bas_mean = 999, bas_drive_mean = 999,
                     bas_funseek_mean = 999, bas_reward_mean = 999,
                     mca_gltfeel_mean = 999,
                     ksads_mdd_anhed = 999)) %>%
  as.mids(.)

write.mice.imputation(mi.res = mi_data_3_mplus, 
                      name = "mi_data_p1_mplus", 
                      dattype = "csv",
                      include.varnames = F,
                      mids2spss = F)
```

