---
title: "10. Correlations Among Composite Scores"
author: "Nicolas L. Camacho"
date: "`r Sys.Date()`"
output: html_document
---

# Install packages (where necessary) and load libraries to run this script
```{r, echo = TRUE, message = FALSE, warning = FALSE}
required_packages <- c("knitr", "tidyverse", "miceadds", "psych", "ggplot2", 
                       "lavaan.mi")
installed <- required_packages %in% rownames(installed.packages())
if(any(!installed)) {
  install.packages(required_packages[!installed])
}
invisible(lapply(required_packages, library, character.only = T))
```

# Set up the root directory from which the cleaned dataset is pulled
```{r setup, echo = TRUE, message = FALSE, warning = FALSE}
opts_knit$set(root.dir = "C:/Users/nicol/Box/01. Nicolas Dissertation/Data/multiply_imputed_data/")
# Update this section to reflect the file path that contains the imputed data
```

# Load multiple imputed data 
```{r}
mi_mplus <- load.Rdata2(filename = "multiply_imputed_data.RData", path = getwd())
```

# Create mean scores for PFC-S General Factor
```{r}
mi_mplus_2 <- complete(mi_mplus, "long", include = T) %>%
  mutate(sex = as.factor(sex)) %>% 
  mutate(across(pfc01:pfc23, as.character)) %>%
  mutate(across(pfc01:pfc23, as.numeric)) %>%
  mutate(pfcs_gen_mean = rowMeans(select(., pfc01:pfc15, pfc17:pfc23), na.rm = T)) %>% 
  as.mids(.)
```

# Polyserial correlation between KSADS Anhedonia item and PFC-S general composite scores
```{r}
pfcs_anh <- 'pfcs_gen_mean ~~ ksads_mdd_anhed'
pfcs_anh_mod <- sem.mi(model = pfcs_anh, data = mi_mplus_2, ordered = "ksads_mdd_anhed")
summary(pfcs_anh_mod, standardized = T)
se_anh <- parameterEstimates.mi(pfcs_anh_mod, standardized = T)$se[1]
cor_anh <- parameterEstimates.mi(pfcs_anh_mod, standardized = T)$std.all[1]
```

# Point-biserial correlation between child sex and PFC-S general composite scores
```{r}
pfcs_sex <- 'pfcs_gen_mean ~~ sex'
pfcs_sex_mod <- sem.mi(model = pfcs_sex, data = mi_mplus_2, ordered = "sex")
summary(pfcs_sex_mod, standardized = T)
se_sex <- parameterEstimates.mi(pfcs_sex_mod, standardized = T)$se[1]
cor_sex <- parameterEstimates.mi(pfcs_sex_mod, standardized = T)$std.all[1]
```

# Correlation table between PFC-S general composite scores and external measures
```{r}
pfcs_gen_corr_df <- micombine.cor(mi_mplus_2, 
                                  variables = c("pfcs_gen_mean", "age_mos",
                                                "bdi_mean", "bis_mean", "bas_mean", 
                                                "bas_drive_mean", "bas_funseek_mean", 
                                                "bas_reward_mean", "erc_er_mean", 
                                                "erc_ln_mean", "mca_gltfeel_mean", 
                                                "brief_gec_t_comb")) %>%
  filter(., str_detect(variable1, "pfcs")) %>%
  dplyr::select(., variable2, r, lower95, upper95) %>% 
  rename(cor_var = variable2) %>% 
  add_row(cor_var = "sex", r = cor_sex, lower95 = (cor_sex - 1.96*se_sex), 
          upper95 = (cor_sex + 1.96*se_anh), .after = 1) %>%
  add_row(cor_var = "ksads_anhedonia", r = cor_anh, lower95 = (cor_anh - 1.96*se_anh), 
          upper95 = (cor_anh + 1.96*se_anh), .after = 12) %>% 
  mutate(., sig = factor(if_else((r > 0 & lower95 > 0) |
                                   (r < 0 & upper95 < 0), 1, 0)))
kable(pfcs_gen_corr_df)
```

``` {r, fig.width = 6, fig.height = 6}
level_order <- c("Age (months)", "Sex", 
                 "KSADS-EC Anhedonia", "MC2 Guilt Feelings", "BRIEF GEC T-Scores",
                 "Caregiver BDI-II", "BIS", "BAS", 
                 "BAS Drive", "BAS Fun Seeking", "BAS Reward Responsiveness", 
                 "ERC Emotion Regulation", "ERC Lability/Negativity")
pfcs_gen_corr_df %>% 
  mutate(cor_var = case_when(cor_var == "age_mos" ~ "Age (months)",
                             cor_var == "sex" ~ "Sex",
                             cor_var == "bdi_mean" ~ "Caregiver BDI-II",
                             cor_var == "bis_mean" ~ "BIS",
                             cor_var == "bas_mean" ~ "BAS",
                             cor_var == "bas_drive_mean" ~ "BAS Drive",
                             cor_var == "bas_funseek_mean" ~ "BAS Fun Seeking",
                             cor_var == "bas_reward_mean" ~ "BAS Reward Responsiveness",
                             cor_var == "erc_er_mean" ~ "ERC Emotion Regulation",
                             cor_var == "erc_ln_mean" ~ "ERC Lability/Negativity",
                             cor_var == "mca_gltfeel_mean" ~ "MC2 Guilt Feelings",
                             cor_var == "brief_gec_t_comb" ~ "BRIEF GEC T-Scores",
                             cor_var == "ksads_anhedonia" ~ "KSADS-EC Anhedonia")) %>%
  ggplot(aes(r, factor(cor_var, levels = level_order), color = sig)) + 
  geom_vline(xintercept = 0, lty = 2, alpha = 0.2) +
  scale_color_manual(values = c("1" = "black", "0" = "darkgray")) +
  geom_point() +
  geom_text(aes(label = sprintf("%.2f", r), vjust = -1),
            family = "serif", size = 4.5) +
  geom_errorbarh(aes(xmin = lower95, xmax = upper95, height = 0.5)) +
  scale_x_continuous(breaks = seq(-1, 1, 0.2)) +
  labs(x = "Correlation", y = "Validation Measures") +
  theme_classic() +
  theme(legend.position = "none",
        text = element_text(family = "serif", size = 15))
```



