---
title: "11. Visualizing Correlations from Mplus"
author: "Nicolas L. Camacho"
date: "`r Sys.Date()`"
output: html_document
---

# Install packages (where necessary) and load libraries to run this script
```{r, echo = TRUE, message = FALSE, warning = FALSE}
required_packages <- c("knitr", "tidyverse", "MplusAutomation", "ggplot2")
installed <- required_packages %in% rownames(installed.packages())
if(any(!installed)) {
  install.packages(required_packages[!installed])
}
invisible(lapply(required_packages, library, character.only = T))
```

# Set up the root directory where the Mplus output files are saved
```{r setup, echo = TRUE, message = FALSE, warning = FALSE}
opts_knit$set(root.dir = "C:/Users/nicol/Box/01. Nicolas Dissertation/Data/Analyses/To Share")
# Update this section to reflect the file path that contains the Mplus output
```

# Load all Mplus output files
```{r}
all_mplus_output <- readModels(target = getwd(), recursive = T)
```

# Pull the necessary details from each output file regarding correlations
# between PFC-S and external measures

## PFC-S, BIS/BAS, ERC, and BDI-II
```{r}
bisbas_erc_bdi <- all_mplus_output$C..Users.nicol.Box.01..Nicolas.Dissertation.Analyses.To.Share.07..correlation.model.1.out
bisbas_erc_bdi_df <- bisbas_erc_bdi$parameters$stdyx.standardized %>%
  filter(str_detect(paramHeader, "LV.WITH") & str_starts(param, "G")) %>% 
  dplyr::select(., paramHeader, est, se) %>% 
  rename(cor_var = paramHeader, r = est) %>% 
  mutate(lower95 = r - (1.96*se), upper95 = r + (1.96*se)) %>% 
  mutate(cor_var = case_when(cor_var == "BDI_LV.WITH" ~ "bdi_mean",
                             cor_var == "BIS_LV.WITH" ~ "bis_mean",
                             cor_var == "BAS_LV.WITH" ~ "bas_mean",
                             cor_var == "DRV_LV.WITH" ~ "bas_drive_mean",
                             cor_var == "FUN_LV.WITH" ~ "bas_funseek_mean",
                             cor_var == "REW_LV.WITH" ~ "bas_reward_mean",
                             cor_var == "ER_LV.WITH" ~ "erc_er_mean",
                             cor_var == "LN_LV.WITH" ~ "erc_ln_mean"))
```

## PFC-S and My Child 2 Guilt Feelings
```{r}
mc2_gltfeel <- all_mplus_output$C..Users.nicol.Box.01..Nicolas.Dissertation.Analyses.To.Share.08a..correlation.model.2...my.child.2.out
mc2_gltfeel_df <- mc2_gltfeel$parameters$stdyx.standardized %>% 
  filter(paramHeader == "G.WITH" & param == "GLT_LV") %>% 
  dplyr::select(., param, est, se) %>% 
  rename(cor_var = param, r = est) %>% 
  mutate(lower95 = r - (1.96*se), upper95 = r + (1.96*se)) %>% 
  mutate(cor_var = "mca_gltfeel_mean")
```

## PFC-S and BRIEF GEC
```{r}
brief_gec <- all_mplus_output$C..Users.nicol.Box.01..Nicolas.Dissertation.Analyses.To.Share.08b..correlation.model.2...brief.gec.out
brief_gec_df <- brief_gec$parameters$stdyx.standardized %>% 
  filter(paramHeader == "G.WITH" & param == "BRIEF_GEC") %>% 
  dplyr::select(., param, est, se) %>% 
  rename(cor_var = param, r = est) %>% 
  mutate(lower95 = r - (1.96*se), upper95 = r + (1.96*se)) %>% 
  mutate(cor_var = "brief_gec_t_comb")
```

## PFC-S and KSADS Anhedonia Item
```{r}
ksads_anh <- all_mplus_output$C..Users.nicol.Box.01..Nicolas.Dissertation.Analyses.To.Share.08c..correlation.model.2...ksads.anhedonia.out
ksads_anh_df <- ksads_anh$parameters$stdyx.standardized %>% 
  filter(paramHeader == "G.WITH" & param == "KS_ANHED") %>% 
  dplyr::select(., param, est, se) %>% 
  rename(cor_var = param, r = est) %>% 
  mutate(lower95 = r - (1.96*se), upper95 = r + (1.96*se)) %>% 
  mutate(cor_var = "ksads_anhedonia")
```

## PFC-S and Child Age
```{r}
child_age <- all_mplus_output$C..Users.nicol.Box.01..Nicolas.Dissertation.Analyses.To.Share.09a..exploratory.correlation.bw.pfc.s.and.child.age.out
child_age_df <- child_age$parameters$stdyx.standardized %>% 
  filter(paramHeader == "G.WITH" & param == "AGE") %>% 
  dplyr::select(., param, est, se) %>% 
  rename(cor_var = param, r = est) %>% 
  mutate(lower95 = r - (1.96*se), upper95 = r + (1.96*se)) %>% 
  mutate(cor_var = "age_mos")
```

## PFC-S and Child Sex
```{r}
child_sex <- all_mplus_output$C..Users.nicol.Box.01..Nicolas.Dissertation.Analyses.To.Share.09b..exploratory.correlation.bw.pfc.s.and.child.sex.out
child_sex_df <- child_sex$parameters$stdyx.standardized %>% 
  filter(paramHeader == "G.WITH" & param == "SEX") %>% 
  dplyr::select(., param, est, se) %>% 
  rename(cor_var = param, r = est) %>% 
  mutate(lower95 = r - (1.96*se), upper95 = r + (1.96*se)) %>%
  mutate(cor_var = "sex")
```

## Combine all correlation data & produce visualization
```{r, fig.width = 6, fig.height = 6}
comb_df <- bisbas_erc_bdi_df %>% 
  bind_rows(mc2_gltfeel_df, brief_gec_df, ksads_anh_df, child_age_df, child_sex_df) %>% 
  mutate(sig = factor(if_else((r > 0 & lower95 > 0) | (r < 0 & upper95 < 0), 1, 0)))

level_order <- c("Age (months)", "Sex", 
                 "KSADS-EC Anhedonia", "MC2 Guilt Feelings", "BRIEF GEC T-Scores", 
                 "Caregiver BDI-II", "BIS", "BAS", 
                 "BAS Drive", "BAS Fun Seeking", "BAS Reward Responsiveness", 
                 "ERC Emotion Regulation", "ERC Lability/Negativity")

comb_df %>% 
  mutate(cor_var = case_when(cor_var == "age_mos" ~ "Age (months)",
                             cor_var == "sex" ~ "Sex",
                             cor_var == "bdi_mean" ~ "Caregiver BDI-II",
                             cor_var == "bis_mean" ~ "BIS",
                             cor_var == "bas_mean" ~ "BAS",
                             cor_var == "bas_drive_mean" ~ "BAS Drive",
                             cor_var == "bas_funseek_mean" ~ "BAS Fun Seeking",
                             cor_var == "bas_reward_mean" ~ "BAS Reward Responsiveness",
                             cor_var == "erc_er_mean" ~ "ERC Emotion Regulation",
                             cor_var == "erc_ln_mean" ~ "ERC Lability/Negativity",
                             cor_var == "mca_gltfeel_mean" ~ "MC2 Guilt Feelings",
                             cor_var == "brief_gec_t_comb" ~ "BRIEF GEC T-Scores",
                             cor_var == "ksads_anhedonia" ~ "KSADS-EC Anhedonia")) %>%
  ggplot(aes(r, factor(cor_var, levels = level_order), color = sig)) + 
  geom_vline(xintercept = 0, lty = 2, alpha = 0.2) +
  scale_color_manual(values = c("1" = "black", "0" = "darkgray")) +
  geom_point() +
  geom_text(aes(label = sprintf("%.2f", r), vjust = -1),
            family = "serif", size = 4.5) +
  geom_errorbarh(aes(xmin = lower95, xmax = upper95, height = 0.5)) +
  scale_x_continuous(breaks = seq(-1, 1, 0.2)) +
  labs(x = "Correlation", y = "Validation Measures") +
  theme_classic() +
  theme(legend.position = "none",
        text = element_text(family = "serif", size = 15))
```

