---
title: "15. Logistic Regressions with PFC-S General Factor Scores"
author: "Nicolas L. Camacho"
date: "2025-08-27"
output: html_document
---

# Install packages (where necessary) and load libraries to run this script
```{r, echo = TRUE, message = FALSE, warning = FALSE}
required_packages <- c("knitr", "tidyverse", "stats")
installed <- required_packages %in% rownames(installed.packages())
if(any(!installed)) {
  install.packages(required_packages[!installed])
}
invisible(lapply(required_packages, library, character.only = T))
```

# Set up the root directory from which the dataset with PFC-S factor scores and CBCL/KSADS-EC based groups is pulled
```{r setup, echo = TRUE, message = FALSE, warning = FALSE}
opts_knit$set(root.dir = "C:/Users/nicol/Box/01. Nicolas Dissertation/Data/")
# Update this section to reflect the file path that contains the correct dataset
```

# Load data and retain only necessary columns
```{r, message = FALSE}
trunc_data <- read_csv("trunc_fscores_grps_updated.csv") %>% 
  dplyr::select(subid, pfc_gen_fs, cbcl_int_grp:cbcl_tot_grp,
                contains("_v_")) %>% 
  mutate(across(c(cbcl_int_grp:anxPlus_noMDD_v_mddPlus_noANX), as.factor))
```

# Run logistic regressions

## KSADS

### No vs. Any Single Diagnosis
```{r}
no_v_anySingle_mod <- glm(no_v_anySingle ~ 1 + pfc_gen_fs, data = trunc_data,
                          family = binomial(link = "logit"))
summary(no_v_anySingle_mod)
confint(no_v_anySingle_mod)
```

### No vs. Any Co-Occurring Diagnoses
```{r}
no_v_anyCO_mod <- glm(no_v_anyCO ~ 1 + pfc_gen_fs, data = trunc_data,
                      family = binomial(link = "logit"))
summary(no_v_anyCO_mod)
confint(no_v_anyCO_mod)
```

### No vs. MDD+/CO-
```{r}
no_v_mddPlus_noCO_mod <- glm(no_v_mddPlus_noCO ~ 1 + pfc_gen_fs, data = trunc_data,
                   family = binomial(link = "logit"))
summary(no_v_mddPlus_noCO_mod)
confint(no_v_mddPlus_noCO_mod)
```

### No vs. MDD+/CO+
```{r}
no_v_mddPlus_coPlus_mod <- glm(no_v_mddPlus_coPlus ~ 1 + pfc_gen_fs, data = trunc_data,
                   family = binomial(link = "logit"))
summary(no_v_mddPlus_coPlus_mod)
confint(no_v_mddPlus_coPlus_mod)
```

### Any/MDD- vs. MDD+/CO-
```{r}
any_noMDD_v_mddPlus_noCO_mod <- glm(any_noMDD_v_mddPlus_noCO ~ 1 + pfc_gen_fs, data = trunc_data,
                   family = binomial(link = "logit"))
summary(any_noMDD_v_mddPlus_noCO_mod)
confint(any_noMDD_v_mddPlus_noCO_mod)
```

### Any/MDD- vs. MDD+/CO+
```{r}
any_noMDD_v_mddPlus_coPlus_mod <- glm(any_noMDD_v_mddPlus_coPlus ~ 1 + pfc_gen_fs, data = trunc_data,
                   family = binomial(link = "logit"))
summary(any_noMDD_v_mddPlus_coPlus_mod)
confint(any_noMDD_v_mddPlus_coPlus_mod)
```

### ANX/MDD- vs. MDD+/ANX-
```{r}
anxPlus_noMDD_v_mddPlus_noANX_mod <- glm(anxPlus_noMDD_v_mddPlus_noANX ~ 1 + pfc_gen_fs, data = trunc_data,
                   family = binomial(link = "logit"))
summary(anxPlus_noMDD_v_mddPlus_noANX_mod)
confint(anxPlus_noMDD_v_mddPlus_noANX_mod)
```

## CBCL Dimensions

### Internalizing
```{r}
int_gen_mod <- glm(cbcl_int_grp ~ 1 + pfc_gen_fs, data = trunc_data,
                   family = binomial(link = "logit"))
summary(int_gen_mod)
confint(int_gen_mod)
```

### Externalizing
```{r}
ext_gen_mod <- glm(cbcl_ext_grp ~ 1 + pfc_gen_fs, data = trunc_data,
                   family = binomial(link = "logit"))
summary(ext_gen_mod)
confint(ext_gen_mod)
```

### Total
```{r}
tot_gen_mod <- glm(cbcl_tot_grp ~ 1 + pfc_gen_fs, data = trunc_data,
                   family = binomial(link = "logit"))
summary(tot_gen_mod)
confint(tot_gen_mod)
```



