---
title: "18a. ROC Curve Analyses with PFC-S General Factor Scores"
author: "Nicolas L. Camacho"
date: "2025-07-31"
output: html_document
---

# Install packages (where necessary) and load libraries to run this script
```{r, echo = TRUE, message = FALSE, warning = FALSE}
required_packages <- c("knitr", "tidyverse", "cutpointr", "emmeans", "ggplot2",
                       "patchwork")
installed <- required_packages %in% rownames(installed.packages())
if(any(!installed)) {
  install.packages(required_packages[!installed])
}
invisible(lapply(required_packages, library, character.only = T))
```

# Set up the root directory from which the dataset with PFC-S factor scores and CBCL/KSADS-EC based groups is pulled
```{r setup, echo = TRUE, message = FALSE, warning = FALSE}
opts_knit$set(root.dir = "C:/Users/nicol/Box/01. Nicolas Dissertation/Data/")
# Update this section to reflect the file path that contains the correct dataset
```

# Load data and retain only necessary columns
```{r, message = FALSE}
trunc_data <- read_csv("trunc_fscores_grps_updated.csv")
```

# KSADS

## Comparison Set 1

### No vs. Any Single Diagnosis
```{r}
# ROC curve metrics
no_v_anySingle_cp <- cutpointr(trunc_data, pfc_gen_fs, no_v_anySingle,
                               method = oc_youden_kernel, metric = youden,
                               pos_class = 1, neg_class = 0, direction = ">=",
                               na.rm = T)
sum_no_v_anySingle_cp <- summary(no_v_anySingle_cp)
sum_no_v_anySingle_cp
# Predicted Probabilities
trunc_2 <- trunc_data %>% mutate(pfc_cat_comp1 = 
                                   case_when(pfc_gen_fs < -0.2477 ~ 0,
                                             pfc_gen_fs >= -0.2477 ~ 1))
no_v_anySingle_mod <- glm(no_v_anySingle ~ 1 + pfc_cat_comp1, data = trunc_2,
                          family = binomial(link="logit"))
emmeans(no_v_anySingle_mod, "pfc_cat_comp1", type = "response")
```

### No vs. Any Co-Occurring Diagnoses
```{r}
# ROC curve metrics
no_v_anyCO_cp <- cutpointr(trunc_data, pfc_gen_fs, no_v_anyCO,
                               method = oc_youden_kernel, metric = youden,
                               pos_class = 1, neg_class = 0, direction = ">=",
                               na.rm = T)
sum_no_v_anyCO_cp <- summary(no_v_anyCO_cp)
sum_no_v_anyCO_cp
# Predicted Probabilities
trunc_2 <- trunc_data %>% mutate(pfc_cat_comp2 = 
                                   case_when(pfc_gen_fs < 0.0323 ~ 0,
                                             pfc_gen_fs >= 0.0323 ~ 1))
no_v_anyCO_mod <- glm(no_v_anyCO ~ 1 + pfc_cat_comp2, data = trunc_2,
                          family = binomial(link="logit"))
emmeans(no_v_anyCO_mod, "pfc_cat_comp2", type = "response")
```

### ROC Curve for Comparison Set 1
```{r, fig.width = 3.5, fig.height = 3.5}
roc_1a <- no_v_anySingle_cp$roc_curve[[1]] %>% mutate(outcome = factor("No vs. Any Single Diagnosis"))
roc_1b <- no_v_anyCO_cp$roc_curve[[1]] %>% mutate(outcome = factor("No vs. Any Co-Occurring Diagnoses"))
roc_1ab <- bind_rows(roc_1a, roc_1b)

p1 <- ggplot(roc_1ab, aes(x = fpr, y = tpr, color = outcome)) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "black") +
  geom_line(linewidth = 0.75) +
  scale_color_manual(values = c("No vs. Any Single Diagnosis" = "#882255", 
                                "No vs. Any Co-Occurring Diagnoses" = "#88CCEE")) +
  annotate("pointrange", x = 1-0.8, y = 0.6552, ymin = 0.6552, ymax = 0.6552,
           size = 0.75, linewidth = 0, colour = "#882255") +
  annotate("pointrange", x = 1-0.8923, y = 0.9212, ymin = 0.9212, ymax = 0.9212,
           size = 0.75, linewidth = 0, colour = "#88CCEE") +
  coord_equal() +
  theme_light(base_family = "serif", base_size = 12) +
  theme(legend.position = "bottom") +
  guides(color = guide_legend(ncol = 1)) +
  labs(x = "1 - Specificity",
       y = "Sensitivity",
       color = NULL)
```


## Comparison Set 2

### No vs. MDD+/CO-
```{r}
# ROC curve metrics
no_v_mddPlus_noCO_cp <- cutpointr(trunc_data, pfc_gen_fs, no_v_mddPlus_noCO,
                                  method = oc_youden_kernel, metric = youden,
                                  pos_class = 1, neg_class = 0, direction = ">=",
                                  na.rm = T)
sum_no_v_mddPlus_noCO_cp <- summary(no_v_mddPlus_noCO_cp)
sum_no_v_mddPlus_noCO_cp
# Predicted Probabilities
trunc_2 <- trunc_data %>% mutate(pfc_cat_comp3 = 
                                   case_when(pfc_gen_fs < -0.221 ~ 0,
                                             pfc_gen_fs >= -0.221 ~ 1))
no_v_mddPlus_noCO_mod <- glm(no_v_mddPlus_noCO ~ 1 + pfc_cat_comp3, data = trunc_2,
                          family = binomial(link="logit"))
emmeans(no_v_mddPlus_noCO_mod, "pfc_cat_comp3", type = "response")
```

### No vs. MDD+/CO+
```{r}
# ROC curve metrics
no_v_mddPlus_coPlus_cp <- cutpointr(trunc_data, pfc_gen_fs, no_v_mddPlus_coPlus,
                                  method = oc_youden_kernel, metric = youden,
                                  pos_class = 1, neg_class = 0, direction = ">=",
                                  na.rm = T)
sum_no_v_mddPlus_coPlus_cp <- summary(no_v_mddPlus_coPlus_cp)
sum_no_v_mddPlus_coPlus_cp
# Predicted Probabilities
trunc_2 <- trunc_data %>% mutate(pfc_cat_comp4 = 
                                   case_when(pfc_gen_fs < 0.0306 ~ 0,
                                             pfc_gen_fs >= 0.0306 ~ 1))
no_v_mddPlus_coPlus_mod <- glm(no_v_mddPlus_coPlus ~ 1 + pfc_cat_comp4, data = trunc_2,
                          family = binomial(link="logit"))
emmeans(no_v_mddPlus_coPlus_mod, "pfc_cat_comp4", type = "response")
```

### ROC Curve for Comparison Set 2
```{r, fig.width = 3.5, fig.height = 3.5}
roc_2a <- no_v_mddPlus_noCO_cp$roc_curve[[1]] %>% mutate(outcome = factor("No vs. MDD/CO-"))
roc_2b <- no_v_mddPlus_coPlus_cp$roc_curve[[1]] %>% mutate(outcome = factor("No vs. MDD/CO+"))
roc_2ab <- bind_rows(roc_2a, roc_2b)

p2 <- ggplot(roc_2ab, aes(x = fpr, y = tpr, color = outcome)) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "black") +
  geom_line(linewidth = 0.75) +
  scale_color_manual(values = c("No vs. MDD/CO-" = "#882255", 
                                "No vs. MDD/CO+" = "#88CCEE")) +
  annotate("pointrange", x = 1-0.8077, y = 0.7619, ymin = 0.7619, ymax = 0.7619,
           size = 0.75, linewidth = 0, colour = "#882255") +
  annotate("pointrange", x = 1-0.8923, y = 0.9394, ymin = 0.9394, ymax = 0.9394,
           size = 0.75, linewidth = 0, colour = "#88CCEE") +
  coord_equal() +
  theme_light(base_family = "serif", base_size = 12) +
  theme(legend.position = "bottom") +
  guides(color = guide_legend(ncol = 1)) +
  labs(x = "1 - Specificity",
       y = "Sensitivity",
       color = NULL)
```

## Comparison Set 3

### Any/MDD- vs. MDD+/CO-
```{r}
# ROC curve metrics
any_noMDD_v_mddPlus_noCO_cp <- cutpointr(trunc_data, pfc_gen_fs, any_noMDD_v_mddPlus_noCO,
                                  method = oc_youden_kernel, metric = youden,
                                  pos_class = 1, neg_class = 0, direction = ">=",
                                  na.rm = T)
sum_any_noMDD_v_mddPlus_noCO_cp <- summary(any_noMDD_v_mddPlus_noCO_cp)
sum_any_noMDD_v_mddPlus_noCO_cp
# Predicted Probabilities
trunc_2 <- trunc_data %>% mutate(pfc_cat_comp5 = 
                                   case_when(pfc_gen_fs < -0.1063 ~ 0,
                                             pfc_gen_fs >= -0.1063 ~ 1))
any_noMDD_v_mddPlus_noCO_mod <- glm(any_noMDD_v_mddPlus_noCO ~ 1 + pfc_cat_comp5, data = trunc_2,
                          family = binomial(link="logit"))
emmeans(any_noMDD_v_mddPlus_noCO_mod, "pfc_cat_comp5", type = "response")
```

### Any/MDD- vs. MDD+/CO+
```{r}
# ROC curve metrics
any_noMDD_v_mddPlus_coPlus_cp <- cutpointr(trunc_data, pfc_gen_fs, any_noMDD_v_mddPlus_coPlus,
                                  method = oc_youden_kernel, metric = youden,
                                  pos_class = 1, neg_class = 0, direction = ">=",
                                  na.rm = T)
sum_any_noMDD_v_mddPlus_coPlus_cp <- summary(any_noMDD_v_mddPlus_coPlus_cp)
sum_any_noMDD_v_mddPlus_coPlus_cp
# Predicted Probabilities
trunc_2 <- trunc_data %>% mutate(pfc_cat_comp6 = 
                                   case_when(pfc_gen_fs < 0.1384 ~ 0,
                                             pfc_gen_fs >= 0.1384 ~ 1))
any_noMDD_v_mddPlus_coPlus_mod <- glm(any_noMDD_v_mddPlus_coPlus ~ 1 + pfc_cat_comp6, data = trunc_2,
                          family = binomial(link="logit"))
emmeans(any_noMDD_v_mddPlus_coPlus_mod, "pfc_cat_comp6", type = "response")
```

### ROC Curve for Comparison Set 3
```{r, fig.width = 3.5, fig.height = 3.5}
roc_3a <- any_noMDD_v_mddPlus_noCO_cp$roc_curve[[1]] %>% mutate(outcome = factor("Any/MDD- vs. MDD/CO-"))
roc_3b <- any_noMDD_v_mddPlus_coPlus_cp$roc_curve[[1]] %>% mutate(outcome = factor("Any/MDD- vs. MDD/CO+"))
roc_3ab <- bind_rows(roc_3a, roc_3b)

p3 <- ggplot(roc_3ab, aes(x = fpr, y = tpr, color = outcome)) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "black") +
  geom_line(linewidth = 0.75) +
  scale_color_manual(values = c("Any/MDD- vs. MDD/CO-" = "#882255", 
                                "Any/MDD- vs. MDD/CO+" = "#88CCEE")) +
  annotate("pointrange", x = 1-0.6452, y = 0.7619, ymin = 0.7619, ymax = 0.7619,
           size = 0.75, linewidth = 0, colour = "#882255") +
  annotate("pointrange", x = 1-0.7742, y = 0.9141, ymin = 0.9141, ymax = 0.9141,
           size = 0.75, linewidth = 0, colour = "#88CCEE") +
  coord_equal() +
  theme_light(base_family = "serif", base_size = 12) +
  theme(legend.position = "bottom") +
  guides(color = guide_legend(ncol = 1)) +
  labs(x = "1 - Specificity",
       y = "Sensitivity",
       color = NULL)
```

## ANX+/MDD- vs. MDD+/ANX-
```{r}
# ROC curve metrics
anxPlus_noMDD_v_mddPlus_noANX_cp <- cutpointr(trunc_data, pfc_gen_fs, 
                                              anxPlus_noMDD_v_mddPlus_noANX,
                                              method = oc_youden_kernel, metric = youden,
                                              pos_class = 1, neg_class = 0, direction = ">=",
                                              na.rm = T)
sum_anxPlus_noMDD_v_mddPlus_noANX_cp <- summary(anxPlus_noMDD_v_mddPlus_noANX_cp)
sum_anxPlus_noMDD_v_mddPlus_noANX_cp
# Predicted Probabilities
trunc_2 <- trunc_data %>% mutate(pfc_cat_comp7 = 
                                   case_when(pfc_gen_fs < 0.1473 ~ 0,
                                             pfc_gen_fs >= 0.1473 ~ 1))
anxPlus_noMDD_v_mddPlus_noANX_mod <- glm(anxPlus_noMDD_v_mddPlus_noANX ~ 1 + pfc_cat_comp7, data = trunc_2,
                          family = binomial(link="logit"))
emmeans(anxPlus_noMDD_v_mddPlus_noANX_mod, "pfc_cat_comp7", type = "response")
```

## ROC Curve for Comparison 7
```{r, fig.width = 3.5, fig.height = 3.5}
roc_4 <- anxPlus_noMDD_v_mddPlus_noANX_cp$roc_curve[[1]] %>% mutate(outcome = factor("ANX/MDD- vs. MDD/ANX-"))

p4 <- ggplot(roc_4, aes(x = fpr, y = tpr, color = outcome)) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "black") +
  geom_line(linewidth = 0.75) +
  scale_color_manual(values = c("black")) +
  annotate("pointrange", x = 1-0.8214, y = 0.8214, ymin = 0.8214, ymax = 0.8214,
           size = 0.75, linewidth = 0, colour = "black") +
  coord_equal() +
  theme_light(base_family = "serif", base_size = 12) +
  theme(legend.position = "bottom") +
  guides(color = guide_legend(ncol = 1)) +
  labs(x = "1 - Specificity",
       y = "Sensitivity",
       color = NULL)
```


```{r, fig.width = 7, fig.height = 7}
(p1 | p2) / (p3 | p4) +
  plot_annotation(tag_levels = "A")
```


# CBCL Dimensions

## Internalizing
```{r}
# ROC curve metrics
int_cp <- cutpointr(trunc_data, pfc_gen_fs, cbcl_int_grp, 
                    method = oc_youden_kernel, metric = youden,
                    pos_class = 1, neg_class = 0, direction = ">=", na.rm = T)
sum_int_cp <- summary(int_cp)
sum_int_cp
# Predicted Probabilities
trunc_2 <- trunc_data %>% mutate(pfc_cat_comp8 = 
                                   case_when(pfc_gen_fs < 0.0847 ~ 0,
                                             pfc_gen_fs >= 0.0847 ~ 1))
int_mod <- glm(cbcl_int_grp ~ 1 + pfc_cat_comp8, data = trunc_2,
               family = binomial(link="logit"))
emmeans(int_mod, "pfc_cat_comp8", type = "response")
```

## Externalizing
```{r}
# ROC curve metrics
ext_cp <- cutpointr(trunc_data, pfc_gen_fs, cbcl_ext_grp, 
                    method = oc_youden_kernel, metric = youden,
                    pos_class = 1, neg_class = 0, direction = ">=", na.rm = T)
sum_ext_cp <- summary(ext_cp)
sum_ext_cp
# Predicted Probabilities
trunc_2 <- trunc_data %>% mutate(pfc_cat_comp9 = 
                                   case_when(pfc_gen_fs < 0.1226 ~ 0,
                                             pfc_gen_fs >= 0.1226 ~ 1))
ext_mod <- glm(cbcl_ext_grp ~ 1 + pfc_cat_comp9, data = trunc_2,
               family = binomial(link="logit"))
emmeans(ext_mod, "pfc_cat_comp9", type = "response")
```

## Total
```{r}
# ROC curve metrics
tot_cp <- cutpointr(trunc_data, pfc_gen_fs, cbcl_tot_grp, 
                    method = oc_youden_kernel, metric = youden,
                    pos_class = 1, neg_class = 0, direction = ">=", na.rm = T)
sum_tot_cp <- summary(tot_cp)
sum_tot_cp
# Predicted Probabilities
trunc_2 <- trunc_data %>% mutate(pfc_cat_comp10 = 
                                   case_when(pfc_gen_fs < 0.1706 ~ 0,
                                             pfc_gen_fs >= 0.1706 ~ 1))
tot_mod <- glm(cbcl_tot_grp ~ 1 + pfc_cat_comp10, data = trunc_2,
               family = binomial(link="logit"))
emmeans(tot_mod, "pfc_cat_comp10", type = "response")
```

## ROC Curve for CBCL Comparison Set
```{r, fig.width = 5, fig.height = 5}
roc_5a <- int_cp$roc_curve[[1]] %>% mutate(outcome = factor("Internalizing"))
roc_5b <- ext_cp$roc_curve[[1]] %>% mutate(outcome = factor("Externalizing"))
roc_5c <- tot_cp$roc_curve[[1]] %>% mutate(outcome = factor("Total Problems"))
roc_5abc <- bind_rows(roc_5a, roc_5b, roc_5c)

ggplot(roc_5abc, aes(x = fpr, y = tpr, color = outcome)) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "black") +
  geom_line(linewidth = 0.75) +
  scale_color_manual(values = c("Internalizing" = "#44AA99", 
                                "Externalizing" = "#CC6677",
                                "Total Problems" = "#332288")) +
  annotate("pointrange", x = 1-0.7076, y = 0.8861, ymin = 0.8861, ymax = 0.8861,
           size = 0.75, linewidth = 0, colour = "#44AA99") +
  annotate("pointrange", x = 1-0.6667, y = 0.8765, ymin = 0.8765, ymax = 0.8765,
           size = 0.75, linewidth = 0, colour = "#CC6677") +
  annotate("pointrange", x = 1-0.7661, y = 0.8798, ymin = 0.8798, ymax = 0.8798,
           size = 0.75, linewidth = 0, colour = "#332288") +
  coord_equal() +
  guides(color = guide_legend(ncol = 3)) +
  labs(x = "1 - Specificity",
       y = "Sensitivity",
       color = NULL) +
  theme_light(base_family = "serif", base_size = 12) +
  theme(legend.position = "bottom")

```
